---
# author:
# - affiliation: 
#   email: 
#   name: Alfonso Garmendia
title: "Data format and survey"
subtitle: "Impact of *Argemone mexicana* L. on tomato plants infected with *Phytophthora infestans*"
title-block-banner: true
date: last-modified
date-format: iso
lang: en-US
code-fold: true
toc: true
toc-location: body
toc-depth: 6
format:
  html:
    embed-resources: true
    self-contained-math: true
---

```{r include=FALSE}
library(readxl)
library(knitr)
library(dplyr)
library(tidyr)
library(writexl)

```

# Original data

```{r download}
### Raw data ---------------------------
data0 <- "data/Data (1).xlsx"

### Download data if not done ----------
if (!file.exists(data0)) 
  source("data-download.R")

```

Original data were downloaded using "*00_data-download.R*" script into "*`r data0`*".

```{r}
s0 <- excel_sheets(data0)
```

The file `r data0` has `r length(s0)` sheets: `r s0`.

## *Phytophthora infestans*

```{r}
#| warning: false
pin0 <- read_excel(data0, s0[1])
```

```{r}
### Variables names --------------------
# names(pin0)
names0 <- c("Treatment", "Block",
  "pi_w1", "pi_w3", "pi_w5",
  "pi_w7", "pi_w9" )
names(pin0) <- names0

### Remove first row -------------------
pin0 <- pin0[-1, ]

```

Variables names are in two rows, therefore:

-   Variables names had to be renamed.
-   First row had to be removed, because is part of the variable names.

The dataframe for the amount of *Phytophthora infestans* (**pin0**) at different weeks is in wide format and has `r length(pin0[[1]])` observations from `r length(pin0)` variables (@tbl-pin0): *`r names(pin0)`*.

```{r tbl-pin0}
#| tbl-cap: Raw data for *Phytophthora infestans*. 
kable(pin0)
```

## Agronomic features

```{r}
# s0
fyi0 <- read_excel(data0, s0[2], 
  range = "A2:I77")
fnu0 <- read_excel(data0, s0[2], 
  range = "K2:S77")
fwe0 <- read_excel(data0, s0[2], 
  range = "U2:X77")
psi0 <- read_excel(data0, s0[2], 
  range = "Z1:AD26")

```

In the sheet for Agronomic features (\*`r s0[2]`) there are four different data tables:

-   *Fruit yield* (**fyi0**) in wide format for different clusters, has `r length(fyi0[[1]])` observations from `r length(fyi0)` variables (@tbl-fyi0): *`r names(fyi0)`*.

-   *Number of fruits* (**fnu0**) in wide format for different clusters, has `r length(fnu0[[1]])` observations from `r length(fnu0)` variables (@tbl-fnu0): *`r names(fnu0)`*.

-   *Average fruits weight* (**fwe0**) has `r length(fwe0[[1]])` observations from `r length(fwe0)` variables (@tbl-fwe0): *`r names(fwe0)`*.

-   *Plant size* (**psi0**) has `r length(psi0[[1]])` observations from `r length(psi0)` variables (@tbl-psi0): *`r names(psi0)`*.

```{r tbl-fyi0}
#| tbl-cap: Raw data for *Fruit yield*. 
kable(fyi0)
```

```{r tbl-fnu0}
#| tbl-cap: Raw data for *Number of fruits*. 
kable(fnu0)
```

```{r tbl-fwe0}
#| tbl-cap: Raw data for *Average fruit weight*. 
kable(fwe0)
```

```{r tbl-psi0}
#| tbl-cap: Raw data for *Plant size*. 
kable(psi0)
```

## Biomarkers

```{r}
bio0 <- read_excel(data0, s0[3])
```

The table for *Biomarkers* (**bio0**) is in long format for different number of days and has `r length(bio0[[1]])` observations from `r length(bio0)` variables (@tbl-bio0): *`r names(bio0)`*.

```{r tbl-bio0}
#| tbl-cap: Raw data for *Biomarkers*. 
kable(bio0)
```

## Non-enzimatic

```{r}
nen0 <- read_excel(data0, s0[4])
```

The table for *non enzimatic* (**nen0**) is in long format for different number of days and has `r length(nen0[[1]])` observations from `r length(nen0)` variables (@tbl-nen0): *`r names(nen0)`*.

```{r tbl-nen0}
#| tbl-cap: Raw data for *non enzimatic*. 
kable(nen0)
```

# Understanding the data format

```{r}
tr <- unique(psi0$Treatment)
# unique(pin0$Treatment)
# unique(fnu0$Treatment)
# unique(fwe0$Treatment)
# unique(fyi0$Treatment)
# unique(bio0$Treatment)
# unique(nen0$Treatment)

```

All data tables have the same `r length(tr)` treatments, but sometimes in different order: *`r tr`*.

```{r}
bl <- unique(psi0$Block)
# unique(pin0$Block)
# unique(fnu0$Block)
# unique(fwe0$Block)
# unique(fyi0$Block)
# unique(bio0$Block)
# unique(nen0$Block)

```

All data tables should have `r length(bl)` blocks (*`r tr`*), but the three fruits tables don't. This will have to be corrected latter.

Therefore, the main structure of data are of 25 observations (plants): 5 blocks x 5 treatments. This structure is the one in *psi0* and *pin0*, but in this last one the measurements are done in 5 different weeks.

The fruits three tables have this same structure times 3 repetitions.

And the markers tables (bio0 and nen0) have this same structure repeated in 5 different times (Day).

# Join tables

I decided to join all fruits and markers tables. There can (and should) be done further combinations for each analysis, but this would be a start. 

For fruits there will be two tables: 
- Table *fruits* with the total values. 
- Table *cluster* with the individual cluster values in long format. 

*Plants* and *Phytophthora* tables are better separate, while *Phytophthora* should be converted to long format.

## Fruits

### Add missing variables

```{r}
b <- numeric()
for (i in 1:5) {
    b <- c(b, rep(i, 3))
}
fnu0$Block <- rep(b, 5)
fyi0$Block <- rep(b, 5)
fwe0$Block <- rep(b, 5)
fwe0$Repetition <- fnu0$Repetition
```

Variables *Block* and *Repetition* were added to the three fruits tables.

### Fruits table

```{r}
# names(fyi0)
n0 <- c("Treatment", "Repetition",
  "Block", "Total")
n1 <- c("Treatment", "Repetition",
  "Block", "tfyield")
fyi1 <- fyi0[n0]
names(fyi1) <- n1

n1 <- c("Treatment", "Repetition",
  "Block", "tfnumber")
fnu1 <- fnu0[n0]
names(fnu1) <- n1

# names(fwe0)
n0 <- c("Treatment", "Repetition",
  "Block", "average weight of fruits")
n1 <- c("Treatment", "Repetition",
  "Block", "fweight")
fwe1 <- fwe0[n0]
names(fwe1) <- n1
```

For each fruit table the Total variable was selected and renamed with a meaningful name:

-   "*tfyield*", for Total fruits yield.
-   "*tfnumber*", for Total fruits number.\
-   "*fweight*", for the average fruits weight.

```{r}
# library(dplyr)
fruits <- left_join(fyi1, fnu1,
  by = join_by(Treatment, Block, 
    Repetition))
fruits <- left_join(fruits, fwe1,
  by = join_by(Treatment, Block, 
    Repetition))
```

A new table *fruits* was created with the totals of fruits variables for each treatment, block and repetition. This table has `r length(fruits[[1]])` observations from `r length(fruits)` variables (@tbl-fruits): *`r names(fruits)`*.

```{r tbl-fruits}
#| tbl-cap: New fruits data with the total data from the three fruits tables. 
kable(fruits)
```

### Long format clusters fruits table

```{r}
# library(tidyr)
w <- which(names(fyi0) == "Total")
fyi1 <- pivot_longer(fyi0[-w], 
  cols = starts_with("cluster"),
  names_to = "Cluster", 
  names_prefix = "cluster ", 
  values_to = "yield")

w <- which(names(fnu0) == "Total")
fnu1 <- pivot_longer(fnu0[-w], 
  cols = starts_with("cluster"),
  names_to = "Cluster", 
  names_prefix = "cluster ", 
  values_to = "fnumber")

cluster <- left_join(fyi1, fnu1,
  by = join_by(Treatment, Block, 
    Repetition, Cluster))

### Clean -------------------------
rm("b", "bl", "data0", "fnu0", "fnu1",
  "fwe0", "fwe1", "fyi0", "fyi1", "i",
  "n0", "n1", "names0", "s0", "tr", 
  "w" )
```

A new table *cluster* was created with the value of yield and fruits number for each cluster,  treatment, block and repetition. This table has `r length(cluster[[1]])` observations from `r length(cluster)` variables (@tbl-cluster): *`r names(cluster)`*.

```{r tbl-cluster}
#| tbl-cap: New cluster data with the yield and number of fruit for each cluster. Only first 20 observations showed. 
kable(cluster[1:20, ])
```

## Markers

```{r}
markers <- left_join(bio0[-4], nen0[-4],
  by = join_by(Day, Treatment, Block))
```

A new table *markers* was created with all the biomarkers and non-enzimatic ones for each day,  treatment and block. This table has `r length(markers[[1]])` observations from `r length(markers)` variables (@tbl-markers): *`r names(markers)`*.

```{r tbl-markers}
#| tbl-cap: New markers data with the biomarkers and non-enzimatic ones. Only first 20 observations showed. 
kable(markers[1:20, ])

rm(bio0, nen0)
```

## *Phytophthora*

```{r}
phytopthora <- pivot_longer(pin0, 
  cols = starts_with("pi_w"),
  names_to = "week", 
  names_prefix = "pi_w", 
  values_to = "Phytopthora")

phytopthora$week <- 
  as.numeric(phytopthora$week)

rm(pin0)
```

A new table *phytopthora* was created with the *Phytopthora* data in each week, treatment and block. This table has `r length(phytopthora[[1]])` observations from `r length(phytopthora)` variables (@tbl-phytopthora): *`r names(phytopthora)`*.

```{r tbl-phytopthora}
#| tbl-cap: New *phytopthora* data for each week in long format. Only first 20 observations showed. 
kable(phytopthora[1:20, ])
```

See in the Article if the weeks correspond to the days in markers table and could then be joined together. 

## Plants

```{r}
plants <- psi0
rm(psi0)
```

The table for plant size is kept as it was. 

# Export data

```{r}
# if (!dir.exists("data")) dir.create("data")
save(fruits, cluster, markers, 
  phytopthora, plants, 
  file = "data/tomato-clean.Rdata")

write_xlsx(list(
  fruits = fruits, 
  cluster = cluster, 
  markers = markers, 
  phytopthora = phytopthora, 
  plants = plants), 
  path = "data/tomato-clean.xlsx")

```

Clean data were exported in two formats inside the "*data/*" folder: 

-   Excel, to check the data visually. 
-   Rdata, to import the data in analysis documents. 

To import the data in further analysis, only the next code is needed: 

`load("data/tomato-clean.Rdata")`

