---
title: "Principal components"
format: html
editor: visual
---

## Principal components graph

Principal Components Analysis (PCA) is a statistical technique used to simplify complex datasets by reducing their dimensionality while retaining as much variability (information) as possible. Biomarkers and Non-enzymatic variables were used for the PCA.

**Required libraries**

```{r, message=FALSE}
library(dplyr)
library(plotly)

```

Call the data

```{r}
library(readxl)
df <- read_excel("DATA/tomato-clean.xlsx", sheet = 3)
```

Since the PCA has to be done with only numerical variables, the numerical variables must be filtered out.

```{r}
df_num <- df %>%
  select(H202, MDA, Superoxide, Phenols, Flavonoids, `Vitamin C`, 
         `Chlorophyll A`, `Chlorophyll B`, `Total chlorophyll`, 
         `B carotene`, Proteins, `yellow carotenoids`)


```

-   If the value is missing (`NA`), not a number (`NaN`), or infinite (`Inf` or `-Inf`), it replaces the value with `0`.

-   Otherwise, it leaves the value unchanged.

```{r}
df_num_clean <- df_num %>%
  mutate_all(~ifelse(is.na(.) | is.nan(.) | is.infinite(.), 0, .))

```

The function prcomp is used to do the principal components analysis.

```{r}
pca_result <- prcomp(df_num_clean, center = TRUE, scale. = TRUE)
summary(pca_result)
```

**`pca_result$x`**: The PCA scores matrix

This contains the transformed data, where each row represents an observation in the new coordinate system (principal components), and each column corresponds to a principal component (e.g., PC1, PC2, etc.).

Adds the `Treatment` and `Day` columns from the original dataset `df` to the PCA scores dataframe. These variables are not part of the PCA but are important for interpreting the results.

```{r}
pca_scores <- as.data.frame(pca_result$x)
pca_scores$Treatment <- df$Treatment  
pca_scores$Day <- df$Day   
```

## Graphic

```{r}
library(ggplot2)
ggplot(pca_scores, aes(x = PC1, y = PC2, color = Treatment, shape = as.factor(Day))) +
  geom_point(size = 3) +
  labs(title = "Principal components Analysis (PCA)",
       x = "(PC1)",
       y = "(PC2)",
       color = "Treatment",
       shape = "Day") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")
```

### Interpretation

#### **Separation Between Treatments:**

The Infes+EXAm and EXAm treatments show a clear differentiation on the right side of the graph, indicating they have a distinct effect on the measured variables compared to Control and Infes.

#### **Control and Infes:**

These two treatments are grouped closely together, suggesting similar behavior in the analyzed variables.

#### **Temporal Evolution:**

As time progresses (especially on days 90 and 105), the separation between treatments becomes more evident, indicating that time influences the observed differences.
