---
title: "Biomarkers"
format: html
editor: visual
---

## Bio markers and Non- enzymatic variables

The data collection started from day 45, where transplantation was carried out. Then every two weeks bio markers and non enzymatic variables were gathered (60, 75, 90, 105 days).

**Required libraries**

```{r,warning=FALSE, message=FALSE}
if (!require("multcompView")) install.packages("multcompView")
if (!require("dplyr")) install.packages("dplyr")
if (!require("PMCMRplus")) install.packages("PMCMRplus")
library(ggplot2)
```

Call the data

```{r}
library(readxl)
df <- read_excel("DATA/tomato-clean.xlsx", sheet = 3)
posthoc_letters_list <- list()
```

## B_carotene

### Statistical analysis

Create an empty list to fill with data of each day.

```{r}
Days <- c(45, 60, 75, 90, 105)

results_table_B_carotene <- data.frame(
  Day = numeric(),
  Test = character(),
  Transformation = character(),
  PostHoc_Method = character(),
  stringsAsFactors = FALSE
)
posthoc_letters_list_B_carotene <- list()

```

In the following for loop iterates over each day in the "Days" variable, performing several statistical operations on the data for each specific day. For each day, it filters the dataset to include only rows corresponding to that day, replaces missing values in the "B carotene" column with zero, and specifies *Block* and *Treatment* as factors. Then, an ANOVA is performed to test the effect of *Treatment* on B carotene*,* and the residuals are checked for normality using the Shapiro-Wilk test. If the residuals are not normally distributed, various transformations (square root, log, inverse) are applied to the data, and the corresponding ANOVA and normality tests are repeated. If all transformations fail to satisfy normality, a non-parametric Friedman test is performed. Post-hoc tests are conducted to identify significant differences between treatments, with the results stored in a table, including the day, test type, transformation used, and post-hoc method.

```{r}
for (dia in Days) {
  # Filtering by day
  df_filtered <- df[df$Day == dia, ]

  # Replace NA values with 0
  df_filtered$B_carotene[is.na(df_filtered$B_carotene)] <- 0

  # Specify factors
  df_filtered$Block <- as.factor(df_filtered$Block)
  df_filtered$Treatment <- as.factor(df_filtered$Treatment)

  # Perform ANOVA, extract residuals, and Shapiro-Wilk test
  anova_result <- aov(B_carotene ~ Treatment, data = df_filtered)
  residuos <- residuals(anova_result)
  shapiro_test_result <- shapiro.test(residuos)

  # Check for normality assumptions
  if (shapiro_test_result$p.value < 0.05) {
    
    # Square root transformation
    df_filtered$B_carotene_sqrt <- sqrt(df_filtered$B_carotene)
    anova_result_sqrt <- aov(B_carotene_sqrt ~ Treatment, data = df_filtered)
    residuos_sqrt <- residuals(anova_result_sqrt)
    shapiro_test_sqrt <- shapiro.test(residuos_sqrt)

    if (shapiro_test_sqrt$p.value < 0.05) {
      
      # Log transformation
      df_filtered$B_carotene_log <- log(df_filtered$B_carotene + 1)
      anova_result_log <- aov(B_carotene_log ~ Treatment, data = df_filtered)
      residuos_log <- residuals(anova_result_log)
      shapiro_test_log <- shapiro.test(residuos_log)

      if (shapiro_test_log$p.value < 0.05) {
        
        # Inverse transformation
        df_filtered$B_carotene_inv <- 1 / (df_filtered$B_carotene + 1)
        anova_result_inv <- aov(B_carotene_inv ~ Treatment, data = df_filtered)
        residuos_inv <- residuals(anova_result_inv)
        shapiro_test_inv <- shapiro.test(residuos_inv)

        if (shapiro_test_inv$p.value < 0.05) {
          
          # Non-parametric test (Friedman)
          friedman_result <- friedman.test(B_carotene ~ Treatment | Block, data = df_filtered)
          
          posthoc <- pairwise.wilcox.test(
            df_filtered$B_carotene,
            df_filtered$Treatment,
            p.adjust.method = "bonferroni"
          )

          letters <- multcompLetters(posthoc$p.value)$Letters
          posthoc_letters_list[[as.character(dia)]] <- letters

          results_table_B_carotene <- rbind(results_table_B_carotene, data.frame(
            Day = dia,
            Test = "Friedman",
            Transformation = "Non-Parametric",
            PostHoc_Method = "Bonferroni"
          ))
        } else {
          
          posthoc <- TukeyHSD(anova_result_inv, "Treatment")
          letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
          posthoc_letters_list[[as.character(dia)]] <- letters

          results_table_B_carotene <- rbind(results_table_B_carotene, data.frame(
            Day = dia,
            Test = "ANOVA",
            Transformation = "Inverse",
            PostHoc_Method = "Tukey"
          ))
        }
      } else {
        
        posthoc <- TukeyHSD(anova_result_log, "Treatment")
        letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
        posthoc_letters_list[[as.character(dia)]] <- letters

        results_table_B_carotene <- rbind(results_table_B_carotene, data.frame(
          Day = dia,
          Test = "ANOVA",
          Transformation = "Log",
          PostHoc_Method = "Tukey"
        ))
      }
    } else {
      
      posthoc <- TukeyHSD(anova_result_sqrt, "Treatment")
      letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
      posthoc_letters_list[[as.character(dia)]] <- letters

      results_table_B_carotene <- rbind(results_table_B_carotene, data.frame(
        Day = dia,
        Test = "ANOVA",
        Transformation = "Square Root",
        PostHoc_Method = "Tukey"
      ))
    }
  } else {
    
    posthoc <- TukeyHSD(anova_result, "Treatment")
    letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
    posthoc_letters_list[[as.character(dia)]] <- letters

    results_table_B_carotene <- rbind(results_table_B_carotene, data.frame(
      Day = dia,
      Test = "ANOVA",
      Transformation = "None",
      PostHoc_Method = "Tukey"
    ))
  }
}

# Remove duplicate rows (if any)
results_table_B_carotene <- unique(results_table_B_carotene)

```

Print the results to know the transformations and the statistical analysis carried out.

```{r}
print(results_table_B_carotene)
```

print the statistical differences for each day

```{r}
for (dia in Days) {
  cat("Día:", dia, "\n")
  print(posthoc_letters_list[[as.character(dia)]])
  cat("\n")
}
```

Graphic

```{r}

```

## Proteins

### Statistical analysis

Create an empty list to fill with data of each day.

```{r}
Days <- c(45, 60, 75, 90, 105)

results_table_Proteins <- data.frame(
  Day = numeric(),
  Test = character(),
  Transformation = character(),
  PostHoc_Method = character(),
  stringsAsFactors = FALSE
)
posthoc_letters_list_Proteins <- list()
```

In the following for loop, it iterates over each day in the "Days" variable, performing several statistical operations on the data for each specific day. For each day, it filters the dataset to include only rows corresponding to that day, replaces missing values in the "Proteins" column with zero, and specifies Block and Treatment as factors. Then, an ANOVA is performed to test the effect of Treatment on Proteins, and the residuals are checked for normality using the Shapiro-Wilk test. If the residuals are not normally distributed, various transformations (square root, log, inverse) are applied to the data, and the corresponding ANOVA and normality tests are repeated. If all transformations fail to satisfy normality, a non-parametric Friedman test is performed. Post-hoc tests are conducted to identify significant differences between treatments, with the results stored in a table, including the day, test type, transformation used, and post-hoc method.

```{r, echo=FALSE}
for (dia in Days) {
  # Filtering by day
  df_filtered <- df[df$Day == dia, ]

  # Replace NA values with 0
  df_filtered$Proteins[is.na(df_filtered$Proteins)] <- 0

  # Specify factors
  df_filtered$Block <- as.factor(df_filtered$Block)
  df_filtered$Treatment <- as.factor(df_filtered$Treatment)

  # Perform ANOVA, extract residuals, and Shapiro-Wilk test
  anova_result <- aov(Proteins ~ Treatment, data = df_filtered)
  residuos <- residuals(anova_result)
  shapiro_test_result <- shapiro.test(residuos)

  # Check for normality assumptions
  if (shapiro_test_result$p.value < 0.05) {
    
    # Square root transformation
    df_filtered$Proteins_sqrt <- sqrt(df_filtered$Proteins)
    anova_result_sqrt <- aov(Proteins_sqrt ~ Treatment, data = df_filtered)
    residuos_sqrt <- residuals(anova_result_sqrt)
    shapiro_test_sqrt <- shapiro.test(residuos_sqrt)

    if (shapiro_test_sqrt$p.value < 0.05) {
      
      # Log transformation
      df_filtered$Proteins_log <- log(df_filtered$Proteins + 1)
      anova_result_log <- aov(Proteins_log ~ Treatment, data = df_filtered)
      residuos_log <- residuals(anova_result_log)
      shapiro_test_log <- shapiro.test(residuos_log)

      if (shapiro_test_log$p.value < 0.05) {
        
        # Inverse transformation
        df_filtered$Proteins_inv <- 1 / (df_filtered$Proteins + 1)
        anova_result_inv <- aov(Proteins_inv ~ Treatment, data = df_filtered)
        residuos_inv <- residuals(anova_result_inv)
        shapiro_test_inv <- shapiro.test(residuos_inv)

        if (shapiro_test_inv$p.value < 0.05) {
          
          # Non-parametric test (Friedman)
          friedman_result <- friedman.test(Proteins ~ Treatment | Block, data = df_filtered)
          
          # Use pairwise Wilcoxon test with exact = FALSE to avoid warning
          posthoc <- pairwise.wilcox.test(
            df_filtered$Proteins,
            df_filtered$Treatment,
            p.adjust.method = "bonferroni",
            exact = FALSE  # Use approximation to avoid tie issue
          )

          letters <- multcompLetters(posthoc$p.value)$Letters
          posthoc_letters_list[[as.character(dia)]] <- letters

          results_table_Proteins <- rbind(results_table_Proteins, data.frame(
            Day = dia,
            Test = "Friedman",
            Transformation = "Non-Parametric",
            PostHoc_Method = "Bonferroni"
          ))
        } else {
          
          posthoc <- TukeyHSD(anova_result_inv, "Treatment")
          letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
          posthoc_letters_list[[as.character(dia)]] <- letters

          results_table_Proteins <- rbind(results_table_Proteins, data.frame(
            Day = dia,
            Test = "ANOVA",
            Transformation = "Inverse",
            PostHoc_Method = "Tukey"
          ))
        }
      } else {
        
        posthoc <- TukeyHSD(anova_result_log, "Treatment")
        letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
        posthoc_letters_list[[as.character(dia)]] <- letters

        results_table_Proteins <- rbind(results_table_Proteins, data.frame(
          Day = dia,
          Test = "ANOVA",
          Transformation = "Log",
          PostHoc_Method = "Tukey"
        ))
      }
    } else {
      
      posthoc <- TukeyHSD(anova_result_sqrt, "Treatment")
      letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
      posthoc_letters_list[[as.character(dia)]] <- letters

      results_table_Proteins <- rbind(results_table_Proteins, data.frame(
        Day = dia,
        Test = "ANOVA",
        Transformation = "Square Root",
        PostHoc_Method = "Tukey"
      ))
    }
  } else {
    
    posthoc <- TukeyHSD(anova_result, "Treatment")
    letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
    posthoc_letters_list[[as.character(dia)]] <- letters

    results_table_Proteins <- rbind(results_table_Proteins, data.frame(
      Day = dia,
      Test = "ANOVA",
      Transformation = "None",
      PostHoc_Method = "Tukey"
    ))
  }
}

# Remove duplicate rows (if any)
results_table_Proteins <- unique(results_table_Proteins)

```

**Analysis of data**

Print the results to know the transformations and the statistical analysis carried out.

```{r}
print(results_table_Proteins)
```

print the statistical differences for each day

```{r}
for (dia in Days) {
  cat("Día:", dia, "\n")
  print(posthoc_letters_list[[as.character(dia)]])
  cat("\n")
}
```

Graphic

```         
```

## Yellow Caratenoids

Repeat the same procedure per variable

```{r, echo=FALSE}
results_table_yellow_carotenoids <- data.frame(
  Day = numeric(),
  Test = character(),
  Transformation = character(),
  PostHoc_Method = character(),
  stringsAsFactors = FALSE
)

# Initialize posthoc letters list
posthoc_letters_list <- list()

```

```{r}
for (dia in Days) {
  cat("Día:", dia, "\n")
  print(posthoc_letters_list[[as.character(dia)]])
  cat("\n")
}

# Loop through each day
for (dia in Days) {
  
  # Filter data for the specific day
  df_filtered <- df[df$Day == dia, ]
  
  # Check if the column 'yellow_carotenoids' exists
  if ("yellow_carotenoids" %in% colnames(df)) {
    
    # Check if the column has NA values and replace them with 0
    if (any(is.na(df_filtered$yellow_carotenoids))) {
      df_filtered$yellow_carotenoids[is.na(df_filtered$yellow_carotenoids)] <- 0
    }
    
    # Ensure 'Block' and 'Treatment' are factors
    df_filtered$Block <- as.factor(df_filtered$Block)
    df_filtered$Treatment <- as.factor(df_filtered$Treatment)
    
    # Perform ANOVA on the original data (yellow_carotenoids ~ Treatment)
    anova_result <- aov(yellow_carotenoids ~ Treatment, data = df_filtered)
    residuos <- residuals(anova_result)
    shapiro_test_result <- shapiro.test(residuos)
    
    # If residuals are not normally distributed, transform the data
    if (shapiro_test_result$p.value < 0.05) {
      
      # Square root transformation
      df_filtered$yellow_carotenoids_sqrt <- sqrt(df_filtered$yellow_carotenoids)
      anova_result_sqrt <- aov(yellow_carotenoids_sqrt ~ Treatment, data = df_filtered)
      residuos_sqrt <- residuals(anova_result_sqrt)
      shapiro_test_sqrt <- shapiro.test(residuos_sqrt)
      
      if (shapiro_test_sqrt$p.value < 0.05) {
       
        # Log transformation
        df_filtered$yellow_carotenoids_log <- log(df_filtered$yellow_carotenoids + 1)  
        anova_result_log <- aov(yellow_carotenoids_log ~ Treatment, data = df_filtered)
        residuos_log <- residuals(anova_result_log)
        shapiro_test_log <- shapiro.test(residuos_log)
        
        if (shapiro_test_log$p.value < 0.05) {
       
          # Friedman test if normality assumption is violated
          friedman_result <- friedman.test(yellow_carotenoids ~ Treatment | Block, data = df_filtered)
          
          # Post-hoc pairwise test using Wilcoxon (Bonferroni adjustment)
          posthoc <- pairwise.wilcox.test(
            df_filtered$yellow_carotenoids,
            df_filtered$Treatment,
            p.adjust.method = "bonferroni"
          )
          
          # Get letters for the post-hoc test
          letters <- multcompLetters(posthoc$p.value)$Letters
          posthoc_letters_list[[as.character(dia)]] <- letters
          
          # Record the results in the results table
          results_table_yellow_carotenoids <- rbind(results_table_yellow_carotenoids, data.frame(
            Day = dia,
            Test = "Friedman",
            Transformation = "Non-Parametric",
            PostHoc_Method = "Bonferroni"
          ))
        } else {
          
          # If normality assumption is met, perform Tukey's test on log-transformed data
          posthoc <- TukeyHSD(anova_result_log, "Treatment")
          letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
          posthoc_letters_list[[as.character(dia)]] <- letters
          
          results_table_yellow_carotenoids <- rbind(results_table_yellow_carotenoids, data.frame(
            Day = dia,
            Test = "ANOVA",
            Transformation = "Log",
            PostHoc_Method = "Tukey"
          ))
        }
      } else {
        
        # If square root transformation works, perform Tukey's test
        posthoc <- TukeyHSD(anova_result_sqrt, "Treatment")
        letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
        posthoc_letters_list[[as.character(dia)]] <- letters
        
        results_table_yellow_carotenoids <- rbind(results_table_yellow_carotenoids, data.frame(
          Day = dia,
          Test = "ANOVA",
          Transformation = "Square Root",
          PostHoc_Method = "Tukey"
        ))
      }
    } else {
      
      # If normality assumption is met, perform Tukey's test on the original data
      posthoc <- TukeyHSD(anova_result, "Treatment")
      letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
      posthoc_letters_list[[as.character(dia)]] <- letters
      
      results_table_yellow_carotenoids <- rbind(results_table_yellow_carotenoids, data.frame(
        Day = dia,
        Test = "ANOVA",
        Transformation = "None",
        PostHoc_Method = "Tukey"
      ))
    }
    
  } else {
    print("Column 'yellow_carotenoids' not found in the data")
  }
}
```

**Analysis of data**

```{r}
print(results_table_Superoxide)

```

```{r}
for (dia in Days) {
  cat("Día:", dia, "\n")
  print(posthoc_letters_list[[as.character(dia)]])
  cat("\n")
}
```

Graph

```{r}
# Initialize results table for yellow carotenoids
results_table_yellow_carotenoids <- data.frame(
  Day = numeric(),
  Test = character(),
  Transformation = character(),
  PostHoc_Method = character(),
  stringsAsFactors = FALSE
)

# Initialize posthoc letters list
posthoc_letters_list <- list()

# Loop through each day
for (dia in Days) {
  
  # Filter data for the specific day
  df_filtered <- df[df$Day == dia, ]
  
  # Check if the column 'yellow_carotenoids' exists
  if ("yellow_carotenoids" %in% colnames(df)) {
    
    # Check if the column has NA values and replace them with 0
    if (any(is.na(df_filtered$yellow_carotenoids))) {
      df_filtered$yellow_carotenoids[is.na(df_filtered$yellow_carotenoids)] <- 0
    }
    
    # Ensure 'Block' and 'Treatment' are factors
    df_filtered$Block <- as.factor(df_filtered$Block)
    df_filtered$Treatment <- as.factor(df_filtered$Treatment)
    
    # Perform ANOVA on the original data (yellow_carotenoids ~ Treatment)
    anova_result <- aov(yellow_carotenoids ~ Treatment, data = df_filtered)
    residuos <- residuals(anova_result)
    shapiro_test_result <- shapiro.test(residuos)
    
    # If residuals are not normally distributed, transform the data
    if (shapiro_test_result$p.value < 0.05) {
      
      # Square root transformation
      df_filtered$yellow_carotenoids_sqrt <- sqrt(df_filtered$yellow_carotenoids)
      anova_result_sqrt <- aov(yellow_carotenoids_sqrt ~ Treatment, data = df_filtered)
      residuos_sqrt <- residuals(anova_result_sqrt)
      shapiro_test_sqrt <- shapiro.test(residuos_sqrt)
      
      if (shapiro_test_sqrt$p.value < 0.05) {
       
        # Log transformation
        df_filtered$yellow_carotenoids_log <- log(df_filtered$yellow_carotenoids + 1)  
        anova_result_log <- aov(yellow_carotenoids_log ~ Treatment, data = df_filtered)
        residuos_log <- residuals(anova_result_log)
        shapiro_test_log <- shapiro.test(residuos_log)
        
        if (shapiro_test_log$p.value < 0.05) {
       
          # Friedman test if normality assumption is violated
          friedman_result <- friedman.test(yellow_carotenoids ~ Treatment | Block, data = df_filtered)
          
          # Post-hoc pairwise test using Wilcoxon (Bonferroni adjustment)
          posthoc <- pairwise.wilcox.test(
            df_filtered$yellow_carotenoids,
            df_filtered$Treatment,
            p.adjust.method = "bonferroni"
          )
          
          # Get letters for the post-hoc test
          letters <- multcompLetters(posthoc$p.value)$Letters
          posthoc_letters_list[[as.character(dia)]] <- letters
          
          # Record the results in the results table
          results_table_yellow_carotenoids <- rbind(results_table_yellow_carotenoids, data.frame(
            Day = dia,
            Test = "Friedman",
            Transformation = "Non-Parametric",
            PostHoc_Method = "Bonferroni"
          ))
        } else {
          
          # If normality assumption is met, perform Tukey's test on log-transformed data
          posthoc <- TukeyHSD(anova_result_log, "Treatment")
          letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
          posthoc_letters_list[[as.character(dia)]] <- letters
          
          results_table_yellow_carotenoids <- rbind(results_table_yellow_carotenoids, data.frame(
            Day = dia,
            Test = "ANOVA",
            Transformation = "Log",
            PostHoc_Method = "Tukey"
          ))
        }
      } else {
        
        # If square root transformation works, perform Tukey's test
        posthoc <- TukeyHSD(anova_result_sqrt, "Treatment")
        letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
        posthoc_letters_list[[as.character(dia)]] <- letters
        
        results_table_yellow_carotenoids <- rbind(results_table_yellow_carotenoids, data.frame(
          Day = dia,
          Test = "ANOVA",
          Transformation = "Square Root",
          PostHoc_Method = "Tukey"
        ))
      }
    } else {
      
      # If normality assumption is met, perform Tukey's test on the original data
      posthoc <- TukeyHSD(anova_result, "Treatment")
      letters <- multcompLetters(posthoc$Treatment[, "p adj"])$Letters
      posthoc_letters_list[[as.character(dia)]] <- letters
      
      results_table_yellow_carotenoids <- rbind(results_table_yellow_carotenoids, data.frame(
        Day = dia,
        Test = "ANOVA",
        Transformation = "None",
        PostHoc_Method = "Tukey"
      ))
    }
    
  } else {
    print("Column 'yellow_carotenoids' not found in the data")
  }
}

# Print posthoc test results for each day
for (dia in Days) {
  cat("Día:", dia, "\n")
  print(posthoc_letters_list[[as.character(dia)]])
  cat("\n")
}

```
