---
title: "Statistical Yield and Number of fruits"
format: html
editor: visual
---

The data is divided into five blocks, six cluster (refers to a group or cluster of tomatoes growing together on the same cluster or stem branch) and five treatments.

**Required libraries**

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(readxl)
library(agricolae)
library(multcompView)
library(dplyr)
library(FSA)
```

Call the data

df is the data taking into account clusters

```{r, message=FALSE}
library(readxl)
df <- read_excel("DATA/tomato-clean.xlsx", sheet = 2)
dftotal <- read_excel("DATA/tomato-clean.xlsx", sheet = 1)
```

Visualize the tables

```{r}
knitr::kable(df, caption = "Data of yield and Number of fruits")
```

In the "Total table" all clusters are summed up.

```{r}
knitr::kable(dftotal, caption = "Data of yield and Number of fruits ")
```

## Fruit Yield

### Statistical analysis

**Analysis of Variance (ANOVA)**

-   Taking into account clusters:

```{r}
Yieldaov <- aov(yield~Treatment*Block, data = df)
summary(Yieldaov)
```

-   Total yield

```{r}
Yieldtotal <- aov(tfyield~Treatment*Block, data = dftotal)
summary(Yieldtotal)
```

**Interpretation:**

Treatments: In both approaches (total and cluster analysis), the treatment factor is significant. This indicates that, regardless of how you group the data (by total treatment or by cluster), treatments have a clear and significant effect on tomato yield.

Blocks: In both analyses, the blocks are not significant. This suggests that, in this case, blocks do not contribute relevant variability to yield, and a block analysis may be unnecessary.

Interaction: In both cases, there is no significant interaction between treatments and blocks, meaning that the effect of treatments is consistent across blocks (whether you consider clusters separately or totals).

**Test of statistical assumptions**

**Normality**

-   Taking into account clusters:

```{r}
resYield <- residuals(Yieldaov)
shapiro.test(resYield)
```

P-value= 2.997e-12 means that the distribution of the data is no normal.

-   Total yield

```{r}
resYieldtotal <- residuals(Yieldtotal)
shapiro.test(resYieldtotal)
```

P-value = 0.1073 means that data has a normal distribution

Since we are interested in knowing the effect of the treatments and not the effect of the clusters, we continue the analysis with the data that contemplates the total, the one that sum of the clusters.

Visualize the behavior of the data

```{r}
qqnorm(resYieldtotal, main = "Q-Q Plot")
qqline(resYieldtotal, col = "red", lwd = 2) 
hist(resYieldtotal,
     main = "Residuals",
     xlab = "Residuals", 
     col = "lightgreen", 
     border = "black",
     breaks = 20)

```

Post hoc test tukey

```{r, warning=FALSE}
TukeyHSD(Yieldtotal, "Treatment")
tukey_result <- TukeyHSD(Yieldtotal, "Treatment")
tukey_pvalues <- tukey_result$Treatment[, "p adj"]
tukey_letters <- multcompLetters(tukey_pvalues)$Letters
tukey_letters
```

**Interpretation:**

EXAm ("a"): These plants belong to group "a," indicating that their yield response is significantly better than infected untreated plants (Infes, "b") and different from the untreated healthy control.

Infes + Captan ("bc"): These plants share a group with Infes ("b") and the Control ("c"), indicating that Captan provides moderate improvement, but not enough to restore their condition to the level of healthy plants treated with EXAm ("a").

Infes + EXAm ("ac"): These plants are in an intermediate group ("ac"), showing results similar to EXAm-treated healthy plants ("a") and also overlapping with group "c" (e.g., Captan-treated or untreated controls).

```{r}
ggplot(dftotal, aes(x = Treatment, y = tfyield)) +
  geom_boxplot(aes(fill = Treatment), color = "black") +  
  geom_text(aes(x = Treatment, 
                y = max(dftotal$tfyield) + 1,  
                label = tukey_letters[as.character(Treatment)]), 
            size = 2) +  
  labs(title = "Effect of treatment on yield",
       x = "Treatment",
       y = "Yield") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Number of fruits

### Statistical analysis

**Analysis of Variance (ANOVA)**

-   Taking into account clusters:

    ```{r}
    fruitnoaov <- aov(fnumber~Treatment*Block, data = df)
    summary(fruitnoaov)
    ```

-   Total Number of fruits

```{r}
fruitnototal <- aov(tfnumber~Treatment*Block, data = dftotal)
summary(fruitnototal)
```

**Interpretation**

Treatment: In the clustered model, Treatment is significant but less strongly (p=0.0181) than in the non-clustered model (p=0.000149). Clustering accounts for intra-group correlations, which increases variability, reducing significance.

Block is consistently non-significant in both models, suggesting that grouping observations into blocks does not contribute much to explaining variability.

Treatment:Block interaction is not significant in either case, indicating treatment effects are consistent across blocks.

**Test of statistical assumptions**

**Normality**

-   Taking into account clusters:

    ```{r}
    resfruitno <- residuals(fruitnoaov)
    shapiro.test(resfruitno)
    ```

    p- value \<0.05 means that is no normal.

-   Total number of fruits

```{r}
resfruitnotot <- residuals(fruitnototal)
shapiro.test(resfruitnotot)
```

The data present a normal distribution.

-   Since we are interested in knowing the effect of the treatments and not the effect of the clusters, we continue the analysis with the data that contemplates the total, the one that sum of the clusters.

    Visualize the behavior of the data

    ```{r}
    qqnorm(resfruitnotot, main = "Q-Q Plot")
    qqline(resfruitnotot, col = "red", lwd = 2) 
    hist(resfruitnotot,
         main = "Residuals",
         xlab = "Residuals", 
         col = "blue", 
         border = "black",
         breaks = 20)
    ```

Homocedasticity

```{r}
bartlett.test(resfruitnotot~dftotal$Treatment)
```

There is no statistically significant evidence to suggest that the variances of the response variable differ across the treatment groups. The assumption of homogeneity of variances is satisfied

Post hoc test tukey

```{r, warning=FALSE}
tukey_result1 <- TukeyHSD(fruitnototal, "Treatment")
tukey_result1
tukey_pvalues1 <- tukey_result1$Treatment[, "p adj"]
tukey_letters1 <- multcompLetters(tukey_pvalues1)$Letters
tukey_letters1
```

EXAm ("a"): These plants belong to group "a," indicating that their yield response is significantly better than infected untreated plants (Infes, "b") and different from the untreated healthy control.

Infes + Captan ("bc"): These plants share a group with Infes ("b") and the Control ("c"), indicating that Captan provides moderate improvement, but not enough to restore their condition to the level of healthy plants treated with EXAm ("a").

Infes + EXAm ("ac"): These plants are in an intermediate group ("ac"), showing results similar to EXAm-treated healthy plants ("a") and also overlapping with group "c" (e.g., Captan-treated or untreated controls).

```{r}
ggplot(dftotal, aes(x = Treatment, y = tfnumber)) +
  geom_boxplot(aes(fill = Treatment), color = "black") +  
  geom_text(aes(x = Treatment, 
                y = max(dftotal$tfnumber) + 1,  
                label = tukey_letters1[as.character(Treatment)]), 
            size = 2) +  
  labs(title = "Effect of treatment on number of fruits",
       x = "Treatment",
       y = "Number of fruits") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Fruit Weight

### Statistical analysis

**Analysis of Variance (ANOVA)**

```{r}
FWeightTotal <- aov(fweight~Treatment*Block, data = dftotal) 
summary(FWeightTotal)
```

### Interpretation:

**Treatments:** The analysis revealed that the treatments (Control, EXAm, Infes, Infes + EXAm, Infes + Captan) had no statistically significant effect on fruit weight (p = 0.189), indicating that observed differences were likely due to random variation rather than the treatments.

**Blocks:** Similarly, the block effect (p = 0.927) and the interaction between treatment and block (p = 0.720) were not significant, suggesting that variability across blocks and their interaction with treatments did not meaningfully impact fruit weight.

**Residuals:** A large portion of variability remained unexplained (Residuals Sum of Squares = 22,763), which might reflect unmeasured factors or experimental limitations.

**In General:** The lack of significant effects could stem from insufficient influence of the treatments, variability in application or measurement, or inadequate statistical power due to only three repetitions per block. To improve future analyses, increasing replication, refining measurement techniques, ensuring consistent application of treatments and inoculation, and exploring additional factors such as environmental conditions are recommended.

**Suggestion:** To explore potential trends in the data, post-hoc tests like Tukey's HSD will be performed even if the ANOVA was not significant.

## Test of Statistical Assumptions

**Normality**

```{r}
resFweightTotal <- residuals(FWeightTotal) 
shapiro.test(resFweightTotal)
```

### **Interpretation**

P-value = 0.000099. The p-value is less than 0.05 meaning that it is not normally distributed. This violates the normality assumption of ANOVA.

## Suggested Remedies

Based on the results, the residuals of the original ANOVA model were not normally distributed, suggesting a need for data transformation or alternative methods.

### **1. Data Transformation**

```{r}
# Log transformation
dftotal$fweight_log <- log(dftotal$fweight)

# Re-run ANOVA with transformed data
FWeightTotal_log <- aov(fweight_log ~ Treatment * Block, data = dftotal)
summary(FWeightTotal_log)

# Test normality of residuals
shapiro.test(residuals(FWeightTotal_log))
```

### **Interpretation:**

The log transformation improved the residuals' normality slightly (W increased from 0.91562 to 0.93677), but they are still not normally distributed, as evidenced by the Shapiro-Wilk p-value of 0.001003. This suggests that the log transformation did not fully address the normality issue. Thus, we will shift to non-parametric alternatives like the Kruskal-Wallis Test.

## **2. Non-Parametric Test**

The transformations do not achieve normality, therefore **Kruskal-Wallis test** which does not require the assumption of normality will be performed.

```{r}
FWeightKW <- kruskal.test(fweight ~ Treatment, data = dftotal)
```

### **Interpretation**

Since the **p-value (0.01373)** is **less than 0.05**, this indicates that there is a statistically significant difference in fruit weight distributions among the treatment groups.

## **Homoscedasticity**

Test for homoscedasticity using Bartlett's test.

```{r}
bartlett.test(fweight ~ Treatment, data = dftotal)
```

Since the **Kruskal-Wallis test** indicates a **statistically significant result** (p-value = 0.01373), the next step is to conduct **post-hoc pairwise comparisons** to identify which treatment groups differ significantly from each other.

## **Visualization of the Data's Behaviour**

Usage of **`qqnorm()`** and **`qqline()`** functions to visualize the normality of the residuals or transformed data.

```{r}
qqnorm(resFweightTotal, main = "Q-Q Plot")
qqline(resFweightTotal, col = "red", lwd = 2) 
hist(resFweightTotal,
     main = "Residuals",
     xlab = "Residuals", 
     col = "maroon", 
     border = "black",
     breaks = 20)
```

## **Non-parametric Post-Hoc Test (Dunn's Test)**

The Kruskal-Wallis test suggests significant differences in fruit weight distributions across the treatments. However, further non-parametric post-hoc analysis is needed to determine which specific treatment groups are different, thus, Dunn's test was performed.

```{r}
dunn_results <- dunnTest(fweight ~ Treatment, data = dftotal, method = "bonferroni")
dunn <- dunn_results$res$P.adj
names(dunn) <- dunn_results$res$Comparison
dunnfinal <- multcompLetters(dunn)
dunn_letters_df <- data.frame(
  Treatment = names(dunnfinal$Letters),
  letters = dunnfinal$Letters
)
dunn_letters_df$Treatment <- gsub(" ", "", dunn_letters_df$Treatment)
library(dplyr)
df0_merged <- left_join(dftotal, dunn_letters_df, by = "Treatment")
```

### Visualization of Data

```{r}
letter_positions <- df0_merged %>%
  group_by(Treatment) %>%
  summarise(max_fweight = max(fweight, na.rm = TRUE)) %>%
  left_join(dunn_letters_df, by = "Treatment")
y_adjust <- 0.5
ggplot(df0_merged, aes(x = Treatment, y = fweight)) +
  geom_boxplot(aes(fill = Treatment), color = "black") +  # Boxplot with color
  geom_text(data = letter_positions,                    # Use precomputed letter positions
            aes(x = Treatment, 
                y = max_fweight + y_adjust, 
                label = letters), 
            size = 2, vjust = 0) +  # Position letters just above the boxplots
  labs(
    title = "Effect of Treatment on Fruit Weight",
    x = "Treatment",
    y = "Fruit Weight"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),             # Center the title
    axis.text.x = element_text(angle = 45, hjust = 1)   # Rotate x-axis labels for clarity
  )
```

### Interpretation

The analysis of fruit weight using the Kruskal-Wallis test confirmed a statistically significant difference among treatment groups; however, Dunn's post-hoc test revealed considerable overlap, as indicated by the similar letters shared in the treatments.

Particularly:

-   **Control group** (no treatment, no inoculation) overlaps with the "ab" group, showing no significant difference from certain treatments.

-   **EXAm** (extract without pathogen) exhibited variable performance, appearing in both the **ac** and **cd** groups, suggesting inconsistent effects.

-   **Infes** (pathogen-inoculated, no treatment) was consistently part of the largest group (**abcd**), indicating it does not significantly differ from most other treatments.

-   **Infes + Captan** (fungicide with inoculation) appeared in both **abcd** and **bd** groups, showing a slight trend toward improved performance but without strong statistical separation.

-   **Infes + EXAm** (extract with inoculation) also grouped within **abcd**, demonstrating no clear differentiation from untreated or pathogen-inoculated groups.

In general, while the analysis detected some differences, the substantial overlap between treatments suggests that their effects on fruit weight are not consistently distinct, with only minor trends observed for **EXAm** and **Infes + Captan**.
